---
title: "BIOL 220 Problem Set 01: Biostats Basics"
subtitle: "Answer Key"
format:
  pdf:
    toc: false
    number-sections: false
    colorlinks: true
---


## Populations, estimates, sources of variation {.unnumbered}

::: {.callout-tip}
## Populations, estimates, sources of variation

1. In one or two sentences, what is the difference between a sample and a population? [1 point]

:::: {.callout-note appearance="simple"}

## Possible answer

A population is the collection of all possible sampling units that we could ever collect data about. We will never be able to measure/collect data about the entire population. Instead, we are only able to collect data about a sample. A sample is a finite subset of all the units from the population

::::

2. If you're trying to study differences in species richness across locations, taxonomic misidentification can occur, what category of error or source of variation is this? [1 point]
    - Bad luck
    - Random variation
    - Scientific 
    - Measurement error

:::: {.callout-note appearance="simple"}

## Answer

Measurement error

::::

3. What/which is an estimate (choose all that apply)? [1 point]
    - A property of a population
    - A statistic calculated from a sample
    - A best guess of the value of a parameter
    - The average body mass of 20 Kōlea birds

:::: {.callout-note appearance="simple"}

## Answer

- A statistic calculated from a sample
- A best guess of the value of a parameter
- The average body mass of 20 Kōlea birds

::::

:::

## Random sampling for a clinical trial {.unnumbered}

Suppose you are designing a clinical trial for a new drug and you have enough budget to afford a trial with 100 participants randomly selected from the total (finite) population of adults living in the fictional country of El Dorado. Assume no one is added or subtracted (dies) from the population during sampling. El Dorado is divided into 10 regions that vary widely in their population size. To randomly select participants, you are given a list of all living adults in El Dorado arranged by Region and alphabetically by surname (last name) within Region like this:

 Region | Name        | ID Number |
--------|-------------|-----------|
 A      | Aaronson, A | 1         |
 $\vdots$ | $\vdots$  | $\vdots$  |
 A      | Zykowski, Z | 65184     |
 B      | Aaronson, A | 65185     |
 $\vdots$ | $\vdots$  | $\vdots$  |
 B      | Zykowski, Z | 187706    |
 C      | Aaronson, A | 187707    |
 $\vdots$ | $\vdots$  | $\vdots$  |
 C      | Zykowski, Z | 245906    |

Each adult also has unique ID Number determined by their Region and Name.

::: {.callout-tip}
## Consider the following four sampling procedures to select participants, then answer questions 4 & 5.

> A. Choose the first 100 individuals in the list </br>
> B. Choose the first 10 individuals from each Region </br>
> C. Use a random number generator to select 10 ID Numbers from each region </br>
> D. Use a random number generator to select 100 ID Numbers from the country population


4. Which of the sampling procedures (A, B, C, or D) is most likely to produce a *biased* estimate of the true population response to the drug? [1 point]

:::: {.callout-note appearance="simple"}

## Answer

A. Choose the first 100 individuals in the list

::::


5. After starting the trial on participants, you learn that a technician accidentally used a nonrandom sampling procedure that is likely to bias your estimate. The technician suggests that you ask for more funding to increase the sample size to 200 while continuing to use the same biased sampling procedure. If you adopt their suggestion, what is most likely true about the resulting estimate from the sample of 200? [1 point]

    - More precise and less biased
    - More precise and equally biased
    - Less precise and less biased
    - Less precise and equally biased

:::: {.callout-note appearance="simple"}

## Answer

More precise and equally biased

::::


:::

## Randomly assign individuals to "treatments" {.unnumbered}

::: {.callout-tip}
## Randomly assign students to project groups.

6. [2 points] As discussed in the Syllabus, the Group Project is a major part of this class. All students in your section will be divided into groups of 4 to 5 students each. For this activity, you will **randomly assign students to groups**. Each student must be assigned to one and only one group. Assume there are 23 students in your section, and you can simply "name" each student "Student 01", "Student 02", ..., "Student 23". You can use any method you want to assign groups, as long as it is truly random. You need to upload your work to show how your randomization procedure works. This could be any of the following: 
    * a picture of a piece of paper showing randomization
    * a Google Sheet/Doc with randomization procedure
    * a short video describing and showing your randomization procedure
    * an *R* script
    * something else, as long as it's clear what you did. 

    **You cannot just send the "names" of students in each group; you must show your randomization procedure.**


:::: {.callout-note appearance="simple"}

## A possible R-based answer

\small

```{r}
# create a vector of numbers 1 to 23 representing students
students <- 1:23

# create a vector of group names 
groups <- c(rep("group A", 5), rep("group B", 5), rep("group C", 5), 
            rep("group D", 4), rep("group E", 4))

# if we line up the student "names" (aka numbers) and the group names in a 
# data.frame, that will tell us which student belongs to which group:

group_memebership <- data.frame(student = students, group = groups)
head(group_memebership)

# but that is *not* random, to make random we need to shuffle the order of
# either the students or the groups; letʻs shuffle the groups

# the sample function used like this will randomly shuffle the order of the
# groups vector
rand_groups <- sample(groups)

group_memebership <- data.frame(student = students, group = rand_groups)
head(group_memebership)

```

::::


:::

## Types of variables {.unnumbered}

Imagine you have a dataset from a different clinical trial that looks like this: 

```{r}
#| echo: false

set.seed(123)
n <- 8
id <- c(letters, as.character(0:9)) |> 
    sample(8, replace = TRUE) |> 
    paste(collapse = "") |> 
    replicate(n = n, expr = _)

set.seed(123)
x <- data.frame(patient_ID = id, 
                disease_status = sample(c("tested positive", "tested negative"), 
                                        n, replace = TRUE),
                treatment = c("control", "low_dose", "medium_dose", "high_dose"), 
                age_in_years = sample(34:64, 8, replace = TRUE))

knitr::kable(x)
```

(Assume the above shows only a small subset of the full data)



::: {.callout-tip}
## Types of variables



7a. Which of the following is a *categorical* variable?

7b. Which of the following is a *numerical* variable?

7c. Which of the following is a *ordinal* variable?

- `patient_ID`
- `disease_status`
- `treatment`
- `age_in_years`


Questions 7a-c are worth [3] points in total.


:::: {.callout-note appearance="simple"}

## Answers

7a. Which of the following is a *categorical* variable?

- `disease_status`
- `patient_ID` is not correct: it is true that `patient_ID` takes discrete values like a categorical variable, but in this situation `patient_ID` is not data, it is just a "book keeping" column to help us organize the data

7b. Which of the following is a *numerical* variable?

- `age_in_years`


7c. Which of the following is a *ordinal* variable?

- `treatment`

::::

:::

